# Week 4 Practical Machine Learning Assignment

## Introduction

The goal of this project is to predict the manner in which they did the exercise. This is the “classe” variable in the training set. I will create a report describing how I have built my model, how I have used cross validation, what I think the expected out of sample error is, and why I have made the choices I did. I will also be using the prediction model to predict 20 different test cases.

## Loading packages and data

```{r,results='hide'}
## Loading packages
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)

## Loading data
training <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
testing  <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))

```

## Cleaning the data

```{r}
## Removing variables with zero variance

non_zero_var <- nearZeroVar(training)
training_data <- training[,-non_zero_var]
testing_data <- testing[,-non_zero_var]

## Removing non numeric variables

training_data <- training_data[,-(1:5)]
testing_data <- testing_data[,-(1:5)]

## Removing variables with NA values

na_val_col <- sapply(training_data, function(x) mean(is.na(x))) > 0.95

training_data <- training_data[,na_val_col == FALSE]
testing_data <- testing_data[,na_val_col == FALSE]
dim(training_data)
dim(testing_data)
```

## Partitioning the data

```{r}
inTrain <- createDataPartition(training_data$classe, p=0.7, list=FALSE)
trainset <- training_data[inTrain,]
testset <- training_data[-inTrain,]
testset$classe <- as.factor(testset$classe) ## Converting the class from character to factor

```

## Random Forest

```{r}
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
RF_modfit <- train(classe ~ ., data = trainset, method = "rf", trControl=controlRF)
RF_prediction <- predict(RF_modfit, testset)
RF_pred_conf <- confusionMatrix(RF_prediction, testset$classe)
RF_pred_conf

```

## Gradient Boosting Model

```{r}
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
GBM_modfit <- train(classe ~ ., data = trainset, method = "gbm", verbose = FALSE,trControl = controlGBM)
GBM_prediction <- predict(GBM_modfit, testset)
GBM_pred_conf <- confusionMatrix(GBM_prediction, testset$classe)
GBM_pred_conf
```

The accuracy of Random Forest model is greater than that of Gradient Boosting Model. So we will be using Random Forest model to answer the questions in the quiz

## Applying the predictor to test data
```{r}
Final_prediction <- predict(RF_modfit,testing_data )
Final_prediction
```